Automated Verification for Real-Time Systems using Implicit Clocks and an Extended Antimirov Algorithm

### 2022-04-15 paper/tv.tex
m

\definecolor{darklavender}{rgb}{0, 0, 0}
\definecolor{backgroundColour}{rgb}{0.95,0.95,0.92}
\definecolor{brickred}{rgb}{0.8, 0.25, 0.33}
\definecolor{red}{rgb}{0.6,0,0} 
\definecolor{blue}{rgb}{0,0,0.6}
\definecolor{green}{rgb}{0,0.8,0}
\definecolor{cyan}{rgb}{0.0,0.6,0.6}
\definecolor{cloudwhite}{rgb}{0.9412, 0.9608, 0.8471} 
\definecolor{davysgrey}{rgb}{0.33, 0.33, 0.33}
\definecolor{deepfuchsia}{rgb}{0.76, 0.33, 0.76}
\definecolor{deeplilac}{rgb}{0.6, 0.33, 0.73}
\definecolor{deepskyblue}{rgb}{0.0, 0.75, 1.0}
\definecolor{indianred}{rgb}{0.8, 0.36, 0.36}
\definecolor{airforceblue}{rgb}{0.36, 0.54, 0.66}
\definecolor{darkred}{rgb}{0.55, 0.0, 0.0}
\usepackage{wrapfig}
\newcommand{\effect}{{\ensuremath{\mathrm{\Phi}}}}
\definecolor{color_pure}{RGB}{3,35,14}
\usepackage{epigraph}
\newcommand{\anyevent}[1]{{\textcolor{darkred}
{{\textbf{\footnotesize #1}}}}}
\newcommand{\anynotevent}[1]{{\textcolor{darkred}
{{\textbf{\footnotesize $\overline{\text{#1}}$}}}}}
\newcommand{\code}[1]{{\tt{\ensuremath{\m{#1}}}}}
\newcommand{\codeme}[1]{{\tt{\ensuremath{#1}}}}
\newcommand{\CONTAIN}{\sqsubseteq}
\newcommand{\underscore}{\textcolor{vividauburn}{\code{\_}}}
\newcommand{\m}{\mathit} 
\newcommand{\lappend}{\mathrel{\texttt{++}}}

@@

  \vspace{-3mm}

@@

  \vspace{-2mm}

@@

\vspace{-2mm}

@@

$[FV\text{-}Seq]$ computes  \code{ \{  \Pi_1, \Theta_1\} } 
from \code{e_1}, 
then further gets  \code{ \{  \Pi_2, \Theta_2 \} } by 
continuously computing the behaviors of \code{e_2}, to be the final state.
$[FV\text{-}Par]$ computes behaviors for  \code{e_1} and \code{e_2} independently,
then parallel merges the effects from these two branches.  
{{\small\begin{gather*}
  \code{\frac{ 
 \vdash \{ \Pi, \Theta\}  e_1  \{  \Pi_1, \Theta_1\} 
\ \  \ 
\vdash \{ \Pi_1, \Theta_1\}  e_2   \{  \Pi_2, \Theta_2 \} 
}{
  \vdash \{  \Pi, \Theta \}\  
  e_1;e_2 \ \{ \Pi_2, \Theta_2
  \}} 
   {[FV\text{-}Seq]}}  
\quad 
  \code{\frac{
 \vdash \{ \Pi, \Theta \}  e_1   \{  \Pi_1, \Theta_1\} 
\ \  \ 
\vdash \{  \Pi, \Theta \}  e_2   \{  \Pi_2, \Theta_2 \} 
    }{
\vdash \{   \Pi, \Theta \}\  e_1 || e_2 \ \{ 
\Pi_1  \wedge \Pi_2 , \Theta_1 || \Theta_2
\}}  
{[FV\text{-}Par]}} 
\end{gather*}}}

The \code{[FV\text{-}If\text{-}Else]} computes an 
over-approximation of the program states, by adding different constrains into 
different branches. \code{\Pi \wedge v} enforces 
\code{v} into the pure constraints of every traces in the state, same 
for \code{\Pi \wedge \neg v }.
{{\small\begin{gather*}
  \code{\frac{ 
    \begin{matrix}
\vdash \{ \Pi \wedge v, \Theta    \} \ e_1 \ \{  \Pi_1, \Theta_2 \} 
\qquad 
\vdash \{ \Pi \wedge \neg v , \Theta  \} \ e_2 \ 
\{  \Pi_2, \Theta_2 \}  
\qquad (v\ is\ local)
\end{matrix}}
  {\vdash \s{ \Pi, \Theta}\  \textbf{if}\ v\ \textbf{then}\ e_1\ 
  \textbf{else}\ e_2 \ \s{ \Pi_1, \Theta_2} \cup \s{\Pi_2, \Theta_2} } \  
  {[FV\text{-}If\text{-}Else\text{-}Local]}} 
  \end{gather*}}}
  %\\
%\code{\frac{ 
%  \begin{matrix}
%\vdash \{ \Pi, \Theta    \} \ e_1 \ \{ \Pi_1, \Theta_2 \} 
%\qquad 
%\vdash \{ \Pi , \Theta  \} \ e_2 \ 
%\{  \Pi_2, \Theta_2 \}  
%\qquad (v\ is\ global)
%\end{matrix}}
%{\vdash \s{ \Pi, \Theta}\  \textbf{if}\ v\ \textbf{then}\ e_1\ 
%\textbf{else}\ e_2 \ \s{\Pi, v. \Theta_1 \vee \Theta_2 }} \  
%{[FV\text{-}If\text{-}Else\text{-}Global]}} 
Rule \code{[FV\text{-}Event]} simply constructs an new event and 
concatenates it to the current program states. 
Rule \code{[FV\text{-}Delay]} creates a trace \code{\epsilon\mysharp t}
, where \code{t} is fresh, and concatenates it to the current program states, 
together with the additional constraint \code{t{=}d}. 
  {{\small\begin{gather*}
  \code{\frac{
    \Theta^\prime {=} \Theta\cdot \anyevent{A}(v, \alpha^*) 
  {\vdash \s{ \Pi, \Theta} \  
  \textbf{event} [\anyevent{A}(v, \alpha^*)] 
   \ \{\Pi, \Theta^\prime \}} \   {[FV\text{-}Event]}} 
   \qquad
  \code{\frac{
    \Theta^\prime {=} \Theta\cdot \epsilon\mysharp t \quad (t\ is\ fresh) 
  {\vdash \s{ \Pi, \Theta} \  
  \textbf{delay} [d] 
   \ \{ \Pi \wedge  t {=}d,  \Theta^\prime \}} \   {[FV\text{-}Delay]}} 
\end{gather*}}}
Rule \code{[FV\text{-}Timeout]} computes the effects from \code{e_1} and \code{e_2}
using the starting state \code{\s{\Pi, \epsilon}}. The final states are the union 
of these effects with their corresponding time variables and arithmetic constrains. 
{{\small\begin{gather*}
  \code{\frac{
    \begin{matrix}
      \vdash \{\Pi, \epsilon \}\ e_1 \ \{  \Pi_1, \Theta_1\} 
      \qquad 
      \vdash \{ \Pi, \epsilon \}\ e_2 \ \{  \Pi_2, \Theta_2 \} 
      \s{\Theta_f, \Pi_f} {=} \s{\Pi_1  \wedge t_1{\leq}d, (\Theta_1\mysharp t)}
      \cup 
  \s{\Pi_2  \wedge t_2{>}d,  (\Theta_2\mysharp t)}   
  \quad (t_1, t_2\ are\ fresh)     
\end{matrix}
  {\vdash \s{ \Pi, \Theta} \  
  e_1\ \textbf{timeout} [d] \  e_2
   \ \s{\Pi_f, \Theta \cdot \Theta_f} } \   {[FV\text{-}Timeout]}} 
\end{gather*}}}
Similar to the previous rule, \code{[FV\text{-}Ddl]} computes the effects 
from \code{e} and adds an upper time-bound to the result. 
{{\small\begin{gather*}
  \code{\frac{
    \begin{matrix}
      \vdash \{ \Pi, \epsilon\}\ e \ \{  \Pi_1, \Theta_1\} 
  \quad (t\ is\ fresh)     
\end{matrix}
  {\vdash \s{ \Pi, \Theta} \  
  \textbf{deadline}  [d] \ e 
   \ \s{\Pi_1  \wedge t{\leq}d, \Theta  \cdot (\Theta_1\mysharp t)} }   
    {[FV\text{-}Ddl]}} 
  \end{gather*}}}
  Rule \code{[FV\text{-}Meth]} initializes the current effects state 
using the declared precondition, accumulates the effects from the method body, 
to obtain \code{ \s{\Pi, \Theta}}; then  it
checks the  inclusion between \code{ \s{\Pi, \Theta}} and the concatenation of 
the precondition and the postcondition\footnote{Here, 
\code{\effect_{post}} only needs to capture the effects from the current method body, excluding the history effects specified in \code{\effect_{pre}}. }.
  {{\small\begin{gather*}
\code{\frac{  \vdash \s{\effect_{pre}}
 \ e \  \{ \Pi, \Theta \} \qquad   \{ \Pi, \Theta \} \CONTAIN  
 \effect_{pre} \cdot \effect_{post}}{  \vdash  \ mn  \ {x^*}\ \{\textbf{req}\ \effect_{pre}\ 
  \textbf{ens}\  \effect_{post}\} \ \{ e \}}   {[FV\text{-}Meth]}}  
  \end{gather*}}}

Rule \code{[FV\text{-}Call]} first checks whether the instantiated 
  precondition of callee, \code{[v^*/x^*]\effect_{pre}}, is satisfied by 
  the current program state. When the check is succeeded, the final states are 
  formed by concatenating the  
  instantiated postcondition to the current states. 
  %, in which we use an auxiliary function \code{rev} to reverse the event sequences of effects.
{{\small\begin{gather*}
\code{\frac{
\begin{matrix}
{\code{  mn  \ {x^*}\ \{\textbf{req}\ \effect_{pre}\ \textbf{ens}\  \effect_{post}\} \ \{ e \} \in \mathcal{P}\ \ }}
\\
\text{\code{ \s{ \Pi, \Theta} \CONTAIN [v^*/x^*] \effect_{pre}   
\qquad   \effect_f =  \s{ \Pi, \Theta} \cdot  [v^*/x^*]\effect_{post}}}
\end{matrix}}
{\vdash \s{ \Pi, \Theta} \ mn({v^*})  \ 
\effect_f} \ {[FV\text{-}Call]}}
\end{gather*}}}

@@

Now, the inclusions are disjunction-free formulas. 
Next we provide the definitions and the key implementations\footnote{As 
the implementations according to basic regular expressions can be found 
in prior work \cite{DBLP:conf/fsttcs/KeilT14}. 
%Here, we focus on presenting 
%the definitions and how do we deal with dependent values in the effects, 
%as the key novelties of this work. 
%More detail implementations can be found in the extended technical report\citeref{Demo}.
} of Nullable, First and Derivative respectively. Intuitively, 
the Nullable function \code{\delta_{{\pi}}(\es)} returns a boolean 
value indicating whether \code{{\pi} \wedge \es} contains the empty 
trace; the First function \code{fst_{{\pi}}( \es)} computes a set of 
 initial \emph{heads}, denoted as \code{h},  of \code{{\pi} \wedge \es}; and the Derivative 
function \code{D^{\pi}_{h}(\es)} computes a next-state 
effects after eliminating the head  \code{h} from the current 
effects \code{{\pi} \wedge \es}.

@@

\begin{definition}[Heads]\label{Heads}
If \code{h} is a head of \code{\pi \wedge \es}, then there exists 
\code{\pi^\prime} and \code{\es^\prime}, 
such that \code{\pi \wedge \es} {=} \code{\pi^\prime \wedge h \cdot \es^\prime}. 
   %We adopt the following naming convention for transition labels: 
A head can be  $t$, denoting a pure time passing; 
  $\ev $, denoting an instant event passing; or 
  $(\ev, t)$, denoting an event passing which takes time \code{t}. 
\end{definition}

\begin{definition}[First]\label{First}
  Given any \code{\effect {=} \pi \wedge \es},
  \code{fst_{{\pi}}(\es)} return a set of heads, be the set of initial elements   
  derivable from effects \code{\pi \wedge \es}, where: 
    { 
   \begin{gather*} 
  \code{fst_{{\pi}}( \bot ) {=} \{\} } \qquad\quad 
  \code{fst_{{\pi}}( \epsilon) {=}  \{\} } \qquad\quad
  \code{fst_{{\pi}}( \ev) {=} \{ \ev \}  }
  \qquad\quad
  \code{fst_{{\pi}}(  \epsilon \mysharp  t) {=}  \{ t \}}
  \qquad\quad
  \code{fst_{{\pi}}( {\es^\star} ) {=} fst_{{\pi}}({ \es})}
   \\
   \code{fst_{{\pi}}(\es \mysharp  t) {=}  \{(\ev, t^\prime) \ | \ \ev \in fst_{\pi}(\es) \}} 
   \ \ \  \emph{($t^\prime$\ is\ fresh)}
  \quad  
  \qquad
  \code{fst_{{\pi}}(  \es_1 {\vee} \es_2) {=} fst_{{\pi}}(  \es_1) \cup fst_{{\pi}}(  \es_2)  }  
  \\
  \code{fst_{{\pi}}( \pi^\prime? ) {=} \{\tau(\pi^\prime); \tau(\neg \pi^\prime)\}  }
      \qquad\qquad\qquad 
     \code{fst_{{\pi}}(  \es_1 {||} \es_2) {=} fst_{{\pi}}(  \es_1) \cup fst_{{\pi}}(  \es_2)  }   
     \\
  {
  \code{fst_{{\pi}}(\es_1 \cdot \es_2) {=} } 
  \begin{cases}
        \code{\code{fst_{{\pi}}(  \es_1)} \cup \code{fst_{{\pi}}(  \es_2)} } \quad   & \code{ if\ \  \delta(\es_1) {=} true}\\
        \code{\code{fst_{{\pi}}(  \es_1)}} & \code{  if\ \  \delta(\es_1) {=} false}
      \end{cases} 
      }  
  \end{gather*}
  }
\end{definition}

@@

\begin{definition}[Partial Derivative]\label{Derivative}
  Given any \code{\effect {=}\pi \wedge  \es}, 
  %and a variable valuation function \code{\env}, 
  the partial derivative \code{D^{\pi}_{h}( \es)} 
   computes the  effects for the 
  left quotient \code{(h)^{\text{-}1}\llbracket \pi \wedge \es 
  \rrbracket}\footnote{Intuitively, the partial derivative refers to the left quotient of a language equation, for example, for REs, \code{x^{{-}1} \llbracket x\cdot y\rrbracket {=} y};  \code{y^{{-}1} \llbracket x\cdot y\rrbracket {=} \bot }; and  \code{y^{{-}1} \llbracket x + y\rrbracket {=} \epsilon}. Here we come up with a new notion of partial derivative for \timedEffects. 
  }. 
  %\footnote{As the reader may notice, our First and Derivative function does not handle the timed sequences constructed by \code{||}, as the synchronous parallel is reduced during the normalization phase.}
  \begin{gather*}
  \code{D^{\pi}_{h}( \bot ) {=}}  \code{False \wedge \bot } 
  \qquad\qquad 
  \code{D^{\pi}_{h}(   \epsilon) {=}}  \code{False \wedge \bot } 
  \qquad\qquad
  \code{D^{\pi}_{h}(  \ev^\prime) } {=} 
  \code{  (\pi \wedge h{=}\ev^\prime) \wedge \epsilon} 
  \\
  {
   \code{\code{D^{\pi}_{h}( \pi^\prime?)} {=}}
  \begin{cases}
        \code{\pi \wedge \pi^\prime? } &\code{if\ h \not\Rightarrow  \pi^\prime}   \\
        \code{\pi \wedge \epsilon} & \code{if\ h \Rightarrow \pi^\prime} 
      \end{cases} 
      }
      \qquad \quad \ \ \ 
  {
   \code{D^{\pi}_{h}( \es_1  \cdot  \es_2) {=} }
  \begin{cases}
        \code{D^{\pi}_{h}( \es_1) \cdot  \es_2\ \vee \  D^{\pi}_{h}( \es_2)}  &\code{if\ \delta_{\pi}(\es_1) {=} true}   \\
        \code{D^{\pi}_{h}( \es_2) \cdot  \es_2} & \code{if\ \delta_{\pi}(\es_1) {=} false} 
      \end{cases} 
      }
  \\
  \code{D^{\pi}_{h}(  {\es^\star}) } {=} \code{ D^{\pi}_{h}(  {\es}) \cdot {\es^\star}}  
  \qquad\quad \ \ \ 
  \code{D^{\pi}_{h}(  \es_1 {\vee} \es_2) } {=} \code{ D^{\pi}_{h}(  \es_1) \vee D^{\pi}_{h}(  \es_2)} 
  \qquad \quad\ \ \ 
  \code{D^{\pi}_{h}( \es_1 {||}\es_2)} {=} \code{\Bar{\Bar{D}}^{\pi}_{h}( \es_1)} || \code{\Bar{\Bar{D}}^{\pi}_{h}( \es_2)} 
   \\
   \code{D^{\pi}_{\ev}(  \es \mysharp t) } {=} {\bigvee}
   \code{ \{ (\pi^\prime \wedge \es^\prime  \mysharp t) \ | \ (\pi^\prime, \es^\prime) \in D^{\pi}_{\ev}(\es) \} } 
\qquad 
   \code{D^{\pi}_{t}(  \es \mysharp t^\prime) } {=} 
   \code{ (\pi \wedge t{+}t^{\prime\prime}{=}t^\prime) \wedge  
   \es \mysharp t^{\prime\prime} \ \ \  {(t^{\prime\prime}\ is\ fresh)} } 
  \end{gather*}
  \end{definition}

@@

\subsection{Rewriting Rules}
\label{InferenceRules}
Given the well-defined auxiliary functions above, we now discuss the key steps and related rewriting rules that we may use in such an effects inclusion proof.

@@

\item 
\textbf{Disprove (Heuristic Refutation).}
\label{Refutation}
This rule is used to disprove the inclusions when the antecedent is nullable, while the consequent is not nullable. Intuitively, the antecedent contains at least one more trace (the empty trace) than the consequent. Therefore, the inclusion is invalid.

\vspace{-4mm}
\input{rules/rule2}

>>>>

\definecolor{darklavender}{rgb}{0, 0, 0}
\definecolor{backgroundColour}{rgb}{0.95,0.95,0.92}
\definecolor{brickred}{rgb}{0.8, 0.25, 0.33}
\definecolor{red}{rgb}{0.6,0,0} 
\definecolor{blue}{rgb}{0,0,0.6}
\definecolor{green}{rgb}{0,0.8,0}
\definecolor{cyan}{rgb}{0.0,0.6,0.6}
\definecolor{cloudwhite}{rgb}{0.9412, 0.9608, 0.8471} 
\definecolor{davysgrey}{rgb}{0.33, 0.33, 0.33}
\definecolor{deepfuchsia}{rgb}{0.76, 0.33, 0.76}
\definecolor{deeplilac}{rgb}{0.6, 0.33, 0.73}
\definecolor{deepskyblue}{rgb}{0.0, 0.75, 1.0}
\definecolor{indianred}{rgb}{0.8, 0.36, 0.36}
\definecolor{airforceblue}{rgb}{0.36, 0.54, 0.66}
\definecolor{darkred}{rgb}{0.55, 0.0, 0.0}
\usepackage{wrapfig}
\newcommand{\effect}{{\ensuremath{\mathrm{\Phi}}}}
\definecolor{color_pure}{RGB}{3,35,14}
\usepackage{epigraph}
\newcommand{\anyevent}[1]{{\textcolor{darkred}
{{\textbf{\small #1}}}}}
\newcommand{\anynotevent}[1]{{\textcolor{darkred}
{{\textbf{\footnotesize $\overline{\text{#1}}$}}}}}
\newcommand{\code}[1]{{\tt{\ensuremath{\m{#1}}}}}
\newcommand{\codeme}[1]{{\tt{\ensuremath{#1}}}}
\newcommand{\CONTAIN}{\sqsubseteq}
\newcommand{\underscore}{\textcolor{vividauburn}{\code{\_}}}
\newcommand{\m}{\mathit} 
\newcommand{\lappend}{\mathrel{\texttt{++}}}

@@

  %\vspace{-3mm}

@@

 % \vspace{-2mm}

@@

%\vspace{-2mm}

@@

\vspace{1mm}
$[FV\text{-}Seq]$ computes  \code{ \{  \Pi_1, \Theta_1\} } 
from \code{e_1}, 
then further gets  \code{ \{  \Pi_2, \Theta_2 \} } by 
continuously computing the behaviors of \code{e_2}, to be the final state.
$[FV\text{-}Par]$ computes behaviors for  \code{e_1} and \code{e_2} independently,
then parallel merges the effects from these two branches.  
{{\small\begin{gather*}
  \code{\frac{ 
 \vdash \{ \Pi, \Theta\}  e_1  \{  \Pi_1, \Theta_1\} 
\ \  \ 
\vdash \{ \Pi_1, \Theta_1\}  e_2   \{  \Pi_2, \Theta_2 \} 
}{
  \vdash \{  \Pi, \Theta \}\  
  e_1;e_2 \ \{ \Pi_2, \Theta_2
  \}} 
   {[FV\text{-}Seq]}}  
\quad 
  \code{\frac{
 \vdash \{ \Pi, \Theta \}  e_1   \{  \Pi_1, \Theta_1\} 
\ \  \ 
\vdash \{  \Pi, \Theta \}  e_2   \{  \Pi_2, \Theta_2 \} 
    }{
\vdash \{   \Pi, \Theta \}\  e_1 || e_2 \ \{ 
\Pi_1  \wedge \Pi_2 , \Theta_1 || \Theta_2
\}}  
{[FV\text{-}Par]}} 
\end{gather*}}}

\vspace{1mm}
The \code{[FV\text{-}If\text{-}Else]} computes an 
over-approximation of the program states, by adding different constrains into 
different branches. \code{\Pi \wedge v} enforces 
\code{v} into the pure constraints of every traces in the state, same 
for \code{\Pi \wedge \neg v }.
{{\small\begin{gather*}
  \code{\frac{ 
    \begin{matrix}
\vdash \{ \Pi \wedge v, \Theta    \} \ e_1 \ \{  \Pi_1, \Theta_2 \} 
\qquad 
\vdash \{ \Pi \wedge \neg v , \Theta  \} \ e_2 \ 
\{  \Pi_2, \Theta_2 \}  
\qquad (v\ is\ local)
\end{matrix}}
  {\vdash \s{ \Pi, \Theta}\  \textbf{if}\ v\ \textbf{then}\ e_1\ 
  \textbf{else}\ e_2 \ \s{ \Pi_1, \Theta_2} \cup \s{\Pi_2, \Theta_2} } \  
  {[FV\text{-}If\text{-}Else\text{-}Local]}} 
  \end{gather*}}}
  %\\
%\code{\frac{ 
%  \begin{matrix}
%\vdash \{ \Pi, \Theta    \} \ e_1 \ \{ \Pi_1, \Theta_2 \} 
%\qquad 
%\vdash \{ \Pi , \Theta  \} \ e_2 \ 
%\{  \Pi_2, \Theta_2 \}  
%\qquad (v\ is\ global)
%\end{matrix}}
%{\vdash \s{ \Pi, \Theta}\  \textbf{if}\ v\ \textbf{then}\ e_1\ 
%\textbf{else}\ e_2 \ \s{\Pi, v. \Theta_1 \vee \Theta_2 }} \  
%{[FV\text{-}If\text{-}Else\text{-}Global]}}

\vspace{1mm}
Rule \code{[FV\text{-}Event]} simply constructs an new event and 
concatenates it to the current program states. 
Rule \code{[FV\text{-}Delay]} creates a trace \code{\epsilon\mysharp t}
, where \code{t} is fresh, and concatenates it to the current program states, 
together with the additional constraint \code{t{=}d}. 
  {{\small\begin{gather*}
  \code{\frac{
    \Theta^\prime {=} \Theta\cdot \anyevent{A}(v, \alpha^*) 
    }
  {\vdash \s{ \Pi, \Theta} \  
  \textbf{event} [\anyevent{A}(v, \alpha^*)] 
   \ \{\Pi, \Theta^\prime \}} \   {[FV\text{-}Event]}} 
   \qquad
  \code{\frac{
    \Theta^\prime {=} \Theta\cdot \epsilon\mysharp t \quad (t\ is\ fresh) 
    }
  {\vdash \s{ \Pi, \Theta} \  
  \textbf{delay} [d] 
   \ \{ \Pi \wedge  t {=}d,  \Theta^\prime \}} \   {[FV\text{-}Delay]}} 
\end{gather*}}}

\vspace{1mm}
Rule \code{[FV\text{-}Timeout]} computes the effects from \code{e_1} and \code{e_2}
using the starting state \code{\s{\Pi, \epsilon}}. The final states are the union 
of these effects with their corresponding time variables and arithmetic constrains. 
{{\small\begin{gather*}
  \code{\frac{
    \begin{matrix}
      \vdash \{\Pi, \epsilon \}\ e_1 \ \{  \Pi_1, \Theta_1\} 
      \qquad 
      \vdash \{ \Pi, \epsilon \}\ e_2 \ \{  \Pi_2, \Theta_2 \} 
      \\
      \s{\Theta_f, \Pi_f} {=} \s{\Pi_1  \wedge t_1{\leq}d, (\Theta_1\mysharp t)}
      \cup 
  \s{\Pi_2  \wedge t_2{>}d,  (\Theta_2\mysharp t)}   
  \quad (t_1, t_2\ are\ fresh)     
\end{matrix}
  }
  {\vdash \s{ \Pi, \Theta} \  
  e_1\ \textbf{timeout} [d] \  e_2
   \ \s{\Pi_f, \Theta \cdot \Theta_f} } \   {[FV\text{-}Timeout]}} 
\end{gather*}}}

\vspace{1mm}
Similar to \code{[FV\text{-}Timeout]}, \code{[FV\text{-}Ddl]} computes the effects 
from \code{e} and adds an upper time-bound to the result. 
{{\small\begin{gather*}
  \code{\frac{
    \begin{matrix}
      \vdash \{ \Pi, \epsilon\}\ e \ \{  \Pi_1, \Theta_1\} 
  \quad (t\ is\ fresh)     
\end{matrix}
  }
  {\vdash \s{ \Pi, \Theta} \  
  \textbf{deadline}  [d] \ e 
   \ \s{\Pi_1  \wedge t{\leq}d, \Theta  \cdot (\Theta_1\mysharp t)} }   
    {[FV\text{-}Ddl]}} 
  \end{gather*}}}

  \vspace{1mm}
  Rule \code{[FV\text{-}Meth]} initializes the current effects state 
using the declared precondition, accumulates the effects from the method body, 
to obtain \code{ \s{\Pi, \Theta}}; then  it
checks the  inclusion between \code{ \s{\Pi, \Theta}} and the concatenation of 
the precondition and the postcondition\footnote{Here, 
\code{\effect_{post}} only needs to capture the effects from the current method body, excluding the history effects specified in \code{\effect_{pre}}. }.
  {{\small\begin{gather*}
\code{\frac{  \vdash \s{\effect_{pre}}
 \ e \  \{ \Pi, \Theta \} \qquad   \{ \Pi, \Theta \} \CONTAIN  
 \effect_{pre} \cdot \effect_{post}}{  \vdash  \ mn  \ {x^*}\ \{\textbf{req}\ \effect_{pre}\ 
  \textbf{ens}\  \effect_{post}\} \ \{ e \}}   {[FV\text{-}Meth]}}  
  \end{gather*}}}

  \vspace{1mm}
Rule \code{[FV\text{-}Call]} first checks whether the instantiated 
  precondition of callee, \code{[v^*/x^*]\effect_{pre}}, is satisfied by 
  the current program state. When the check is succeeded, the final states are 
  formed by concatenating the  
  instantiated postcondition to the current states. 
  %, in which we use an auxiliary function \code{rev} to reverse the event sequences of effects.
{{\small\begin{gather*}
\code{\frac{
\begin{matrix}
{\code{  mn  \ {x^*}\ \{\textbf{req}\ \effect_{pre}\ \textbf{ens}\  \effect_{post}\} \ \{ e \} \in \mathcal{P}\ \ }}
\\
\text{\code{ \s{ \Pi, \Theta} \CONTAIN [v^*/x^*] \effect_{pre}   
\qquad   \effect_f =  \s{ \Pi, \Theta} \cdot  [v^*/x^*]\effect_{post}}}
\end{matrix}}
{\vdash \s{ \Pi, \Theta} \ mn({v^*})  \ 
\effect_f} \ {[FV\text{-}Call]}}
\end{gather*}}}

@@

\vspace{2mm}
Now, the inclusions are disjunction-free formulas. 
Next we provide the definitions and the key implementations\footnote{As 
the implementations according to basic regular expressions can be found 
in prior work \cite{DBLP:conf/fsttcs/KeilT14}. 
%Here, we focus on presenting 
%the definitions and how do we deal with dependent values in the effects, 
%as the key novelties of this work. 
%More detail implementations can be found in the extended technical report\citeref{Demo}.
} of Nullable, First and Derivative respectively. Intuitively, 
the Nullable function \code{\delta_{{\pi}}(\es)} returns a boolean 
value indicating whether \code{{\pi} \wedge \es} contains the empty 
trace; the First function \code{fst_{{\pi}}( \es)} computes a set of 
 initial \emph{heads}, denoted as \code{h},  of \code{{\pi} \wedge \es}; and the Derivative 
function \code{D^{\pi}_{h}(\es)} computes a next-state 
effects after eliminating the head  \code{h} from the current 
effects \code{{\pi} \wedge \es}.

@@

\vspace{2mm}
\begin{definition}[Heads]\label{Heads}
If \code{h} is a head of \code{\pi \wedge \es}, then there exists 
\code{\pi^\prime} and \code{\es^\prime}, 
such that \code{\pi \wedge \es} {=} \code{\pi^\prime \wedge h \cdot \es^\prime}. 
   %We adopt the following naming convention for transition labels: 
A head can be  $t$, denoting a pure time passing; 
  $\ev $, denoting an instant event passing; or 
  $(\ev, t)$, denoting an event passing which takes time \code{t}. 
\end{definition}

\vspace{2mm}
\begin{definition}[First]\label{First}
  Given any \code{\effect {=} \pi \wedge \es},
  \code{fst_{{\pi}}(\es)} return a set of heads, be the set of initial elements   
  derivable from effects \code{\pi \wedge \es}, where: 
    { 
   \begin{gather*} 
  \code{fst_{{\pi}}( \bot ) {=} \{\} } \qquad\quad 
  \code{fst_{{\pi}}( \epsilon) {=}  \{\} } \qquad\quad
  \code{fst_{{\pi}}( \ev) {=} \{ \ev \}  }
  \qquad\quad
  \code{fst_{{\pi}}(  \epsilon \mysharp  t) {=}  \{ t \}}
  \qquad\quad
  \code{fst_{{\pi}}( {\es^\star} ) {=} fst_{{\pi}}({ \es})}
   \\
   \code{fst_{{\pi}}(\es \mysharp  t) {=}  \{(\ev, t^\prime) \ | \ \ev \in fst_{\pi}(\es) \}} 
   \ \ \  \emph{($t^\prime$\ is\ fresh)}
  \quad  
  \qquad
  \code{fst_{{\pi}}(  \es_1 {\vee} \es_2) {=} fst_{{\pi}}(  \es_1) \cup fst_{{\pi}}(  \es_2)  }  
  \\
  \code{fst_{{\pi}}( \pi^\prime? ) {=} \{\tau(\pi^\prime); \tau(\neg \pi^\prime)\}  }
      \qquad\qquad\qquad 
     \code{fst_{{\pi}}(  \es_1 {||} \es_2) {=} fst_{{\pi}}(  \es_1) \cup fst_{{\pi}}(  \es_2)  }   
     \\
  {
  \code{fst_{{\pi}}(\es_1 \cdot \es_2) {=} } 
  \begin{cases}
        \code{\code{fst_{{\pi}}(  \es_1)} \cup \code{fst_{{\pi}}(  \es_2)} } \quad   & \code{ if\ \  \delta(\es_1) {=} true}\\
        \code{\code{fst_{{\pi}}(  \es_1)}} & \code{  if\ \  \delta(\es_1) {=} false}
      \end{cases} 
      }  
  \end{gather*}
  }
\end{definition}

@@

\vspace{2mm}
\begin{definition}[Partial Derivative]\label{Derivative}
  Given any \code{\effect {=}\pi \wedge  \es}, 
  %and a variable valuation function \code{\env}, 
  the partial derivative \code{D^{\pi}_{h}( \es)} 
   computes the  effects for the 
  left quotient \code{(h)^{\text{-}1}\llbracket \pi \wedge \es 
  \rrbracket}\footnote{Intuitively, the partial derivative refers to the left quotient of a language equation, for example, for REs, \code{x^{{-}1} \llbracket x\cdot y\rrbracket {=} y};  \code{y^{{-}1} \llbracket x\cdot y\rrbracket {=} \bot }; and  \code{y^{{-}1} \llbracket x + y\rrbracket {=} \epsilon}. Here we come up with a new notion of partial derivative for \timedEffects. 
  }. 
  %\footnote{As the reader may notice, our First and Derivative function does not handle the timed sequences constructed by \code{||}, as the synchronous parallel is reduced during the normalization phase.}
  \begin{gather*}
  \code{D^{\pi}_{h}( \bot ) {=}}  \code{False \wedge \bot } 
  \qquad\qquad 
  \code{D^{\pi}_{h}(   \epsilon) {=}}  \code{False \wedge \bot } 
  \qquad\qquad
  \code{D^{\pi}_{h}(  \ev^\prime) } {=} 
  \code{  (\pi \wedge h{=}\ev^\prime) \wedge \epsilon} 
  \\
  {
   \code{\code{D^{\pi}_{h}( \pi^\prime?)} {=}}
  \begin{cases}
        \code{\pi \wedge \pi^\prime? } &\code{if\ h \not\Rightarrow  \pi^\prime}   \\
        \code{\pi \wedge \epsilon} & \code{if\ h \Rightarrow \pi^\prime} 
      \end{cases} 
      }
      \qquad \quad \ \ \ 
  {
   \code{D^{\pi}_{h}( \es_1  \cdot  \es_2) {=} }
  \begin{cases}
        \code{D^{\pi}_{h}( \es_1) \cdot  \es_2\ \vee \  D^{\pi}_{h}( \es_2)}  &\code{if\ \delta_{\pi}(\es_1) {=} true}   \\
        \code{D^{\pi}_{h}( \es_2) \cdot  \es_2} & \code{if\ \delta_{\pi}(\es_1) {=} false} 
      \end{cases} 
      }
  \\
  \code{D^{\pi}_{h}(  {\es^\star}) } {=} \code{ D^{\pi}_{h}(  {\es}) \cdot {\es^\star}}  
  \qquad\quad \ \ \ 
  \code{D^{\pi}_{h}(  \es_1 {\vee} \es_2) } {=} \code{ D^{\pi}_{h}(  \es_1) \vee D^{\pi}_{h}(  \es_2)} 
  \qquad \quad\ \ \ 
  \code{D^{\pi}_{h}( \es_1 {||}\es_2)} {=} \code{\Bar{\Bar{D}}^{\pi}_{h}( \es_1)} || \code{\Bar{\Bar{D}}^{\pi}_{h}( \es_2)} 
   \\
   \code{D^{\pi}_{\ev}(  \es \mysharp t) } {=} {\bigvee}
   \code{ \{ (\pi^\prime \wedge \es^\prime  \mysharp t) \ | \ (\pi^\prime, \es^\prime) \in D^{\pi}_{\ev}(\es) \} } 
\qquad 
   \code{D^{\pi}_{t}(  \es \mysharp t^\prime) } {=} 
   \code{ (\pi \wedge t{+}t^{\prime\prime}{=}t^\prime) \wedge  
   \es \mysharp t^{\prime\prime} \ \ \  {(t^{\prime\prime}\ is\ fresh)} } 
  \end{gather*}
  \end{definition}

@@

  \vspace{2mm}
\subsection{Rewriting Rules}
\label{InferenceRules}
Given the well-defined auxiliary functions above, we now discuss the key steps and related rewriting rules that we may use in such an effects inclusion proof.

@@

\item 
\textbf{Disprove (Heuristic Refutation).}
\label{Refutation}
This rule is used to disprove the inclusions when the antecedent is nullable, while the consequent is not nullable. Intuitively, the antecedent contains at least one more trace (the empty trace) than the consequent. Therefore, the inclusion is invalid. 
\input{rules/rule2}
